Constraining Black Hole Growth with High-z Quasar Observations

### 2022-09-27 ms.tex
version0926; -- real one; w/ figs folder

\shorttitle{BH growth toward $z=$ 6 BHMF \& QLF}
 \shortauthors{Li et al.}
%%
%% You can add a light gray and diagonal water-mark to the first page 
%% with this command:
% \watermark{DRAFT}
%% where "text", e.g. DRAFT, is the text to appear.  If the text is 
%% long you can control the water-mark size with:
%% \setwatermarkfontsize{dimension}
%% where dimension is any recognized LaTeX dimension, e.g. pt, in, etc.
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\graphicspath{{./}{../figs/}}
% \graphicspath{{./}{figures/}{figs/}}
%% This is the end of the preamble.  Indicate the beginning of the
%% manuscript itself with \begin{document}.

@@

%%%%%%%
%    Fig. 1
%%%%%%%
\begin{figure*}
\centering
\includegraphics[width=160mm]{seedBHMF_z.png}
\caption{
Mass distribution of seed BHs formed in quasar host galaxies at different redshifts ($50\lesssim z \lesssim 10$).
%Each panel includes a redshift span of 10, 
In each panel, the cumulative number distribution of BHs formed prior than the redshift interval is shown with green bars,
and the contribution from newly-born BHs during the interval are highlighted in magenta.
%with the newly formed BHs denoted by the hatched regions, 
%and the cumulative BHs formed previously by the blue region.
The majority of seed BHs form in the epoch of $20\lesssim z \lesssim 40$.
%The most vigorous seed formation epoches are in redshift ranges $z\sim$ $40-30$ and $30-20$.
Overall, the seed BH mass distribution is extended to $M_\bullet \gtrsim 10^5~\Msun$, 
imprinted with various pathways of parent halo assembly and environmental effects as discussed in \S\ref{sec:seed}.
}
\label{fig:seedmf}
\end{figure*}

@@

We describe how to calculate the time evolution of the BH mass function, 
combining the semi-analytical growth model and ERDF (see Eqs.~\ref{eq:mdot} and \ref{eq:Pl}).
The schematic illustration of our numerical procedure is shown in Fig.~\ref{fig:scheme}. 
We first set a series of checkpoints with an interval of $\tlife$, 
tracing back from $t_{\rm H}(z=6)$ to just prior than the first seed formation time $t_1$, where $t_i$ is the $i$-th seed forming time
as defined in \S\ref{sec:seed} ($1\leq i\leq N_{\rm tot}$).
Note that the number of the checkpoints is given by $\lceil (t_{\rm H}-t_1)/\tau \rceil+1$, where $\lceil x\rceil \equiv {\rm min}\{n \in \mathbb{Z}~ |~x\leq n\}$.

@@

For the two cases, the best-fit solution consistently shows a small value of $\delta < 0.1$.
This result favors the nearly exponential BH growth model with minor mass dependence (Eqs.~\ref{eq:mdot} and \ref{eq:f_M}).
The typical duration of quasar activity is required to be $\tlife \simeq 20-30$ Myr for both the cases.
This suggests that accreting seed BHs vary their growth speeds and their Eddington ratios 
$\simeq 30-40$ times on average toward the cosmic time at $z\simeq 6$.
The timescale of $\tlife$ is also related to the quasar lifetimes of $t_{\rm Q}\sim 1-10$ Myr, 
which are estimated by the measurement of the physical extents of hydrogen Ly$\alpha$ proximity zones 
observed in the rest-frame UV spectra \citep[e.g.,][]{2018ApJ...867...30E,2019ApJ...884L..19D}.
More detailed comparison is discussed in \S\ref{sec:evol}.

@@

%%%%%%%
%  Table. 2
%%%%%%%
\begin{deluxetable*}{ccccccccccc}%[t!]
\tablecolumns{1}
\renewcommand\thetable{2} %! fix indexing
\tablewidth{0pt}
\tablecaption{Parametric Quasar Luminosity Function}
\tablehead{
  & & \multicolumn{4}{c}{Unobscured QLF} && \multicolumn{4}{c}{Intrinsic QLF}  \\
  \cline{3-6} \cline{8-11}
  \colhead{Redshift} && \colhead{$\Phi^\ast$} & \colhead{$\Muv^\ast$} & \colhead{$\tilde \alpha$} & \colhead{$ \tilde \beta$}& &
  \colhead{$\Phi^\ast$} & \colhead{$\Muv^\ast$} & \colhead{$\tilde \alpha$} & \colhead{$\tilde \beta$}\\
  & & \colhead{(Gpc$^{-3}$ mag$^{-1}$)} & \colhead{(mag)} & & & &
  \colhead{(Gpc$^{-3}$ mag$^{-1}$)} & \colhead{(mag)} & &
}
\startdata
$z\sim 6$  && 46.5 & -23.62 & -1.33 & -2.57    && 89.2 & -23.73 & -1.59 & -2.67 \\
$z\sim 7$  && 6.98 & -23.70 & -1.61 & -2.82   && 12.4 & -23.85 & -1.87 & -2.92 \\
$z\sim 8$  && 1.18 & -23.80 & -1.83 & -3.02   && 1.91 & -23.98 & -2.08 & -3.12 \\
$z\sim 9$  && 0.268 & -23.80 & -1.98 & -3.18  && 0.429& -23.96 & -2.23 & -3.28 \\
$z\sim 10$ && 0.0841 & -23.68 & -2.01 & -3.30 && 0.144 & -23.82 & -2.35 & -3.40 \\
\enddata
\label{tab:QLF_fit}
\tablecomments{Parametric fitting function of both unobscured and intrinsic QLFs at $z=6-10$, produced by the best-fit model with $\fseed=0.01$.
}
\end{deluxetable*}

%%%%%%%
%    Table. 3
%%%%%%%
\begin{deluxetable*}{lccccccc}%[t!]
\tablecolumns{1}
\renewcommand\thetable{3} %! fix indexing
\tablewidth{0pt} 
\tablecaption{\blue{Detection Number of Quasars Estimated for Future Surveys} \label{tab:N_detect}} %! reference indexing
\tablehead{
  \colhead{Telescope} & \colhead{ Survey area} & \colhead{Photo. depth} &  \colhead{$N(z\sim 6)$} & \colhead{$N(z\sim 7)$}& \colhead{$N(z\sim 8)$}&  \colhead{$N(z\sim 9)$}&  \colhead{$N(z\sim 10)$}\\
  \colhead{} & \colhead{(\rm{deg}$^2$)} & \colhead{(AB mag)} & & & & & 
}
\startdata
Wide RST    &  2000  & $YJH$  27   & $6759_{-3185}^{+15} $ &  $1514_{-558}^{+169}$ &  $349_{-131}^{53}$ & $85_{-48}^{+23}$ & $25^{+12}_{-15}$  \\
Wide RST    &  2000  & $\Muv=-26$ & $21_{-5}^{+13} $      &  $2_{-7}^{+8}$        &  --  & -- & --  \\
Wide Euclid & 15000  & $YJH$  24   & $6654^{+95}_{-2282}$  &  $708^{+213}_{-237}$ &   $93^{+66}_{-56}$ & $14^{+32}_{-27}$ & --  \\
Wide Euclid & 15000  & $YJH$  23   & $2480^{+188}_{-614}$  &  $207^{+144}_{-114}$ &   $23^{+27}_{-19}$ & $3^{+12}_{-10}$ & --  \\
Wide Euclid & 15000  & $\Muv=-26$ & $154^{+100}_{-38}$  &  $12^{+59}_{-50}$ &   $1^{+4}_{-3}$ & -- & --  \\
\enddata
\tablecomments{
Predicted number of quasars at each redshift in the two surveys, with the assumption of the survey volume with $\Delta z =1$ and 100\% completeness.
The integration of QLF starts from the 5$\sigma$ point source detection limits of the $YJH$ bands in each survey, 
and the bright magnitude with $\Muv=-26$ mag.
In addition, detection limit with 22 mag in the $YJH$ bands is examined for Euclid,
mimicking the effect of quasar selection function.% following \citet{2019A&A...631A..85E}.
}
\end{deluxetable*}

@@

%%%%%%%
%    Fig. 7
%%%%%%%
\begin{figure*}[h!]
\centering
\includegraphics[width=85mm]{LF_spreadz7.pdf}\hspace{2mm}
\includegraphics[width=85mm]{LF_spreadz8.pdf}\\\vspace{5mm}
\includegraphics[width=85mm]{LF_spreadz9.pdf}\hspace{2mm}
\includegraphics[width=85mm]{LF_spreadz10.pdf}
\caption{Predicted quasar luminosity functions and depths and volumes of upcoming surveys.
The black solid and dashed curves present the QLF for the unobscured population (e.g., type-I AGNs) 
and the whole population (i.e., type-I and type-II AGNs), and the shaded region presents the $1\sigma$ spread.
The expected depths and volumes of future surveys with Wide RST and Wide Euclid are overlaid with the blue lines,
indicating the detectability of quasars at these redshifts.
Galaxy UV luminosity functions are plotted with the data points (grey) and fitting functions with a Schechter (dotted curve) and double power-law (dash-dotted curve) shape.
The galaxy luminosity function data at $z\sim$ 7 are taken from \citet{2022ApJS..259...20H} and \citet{2021AJ....162...47B}. 
At $z\sim$ 8, the data points are from \citet{2021AJ....162...47B},
\citet{2022arXiv220712356D}, \citet{2013MNRAS.432.2696M}, \citet{2020MNRAS.493.2059B}, and \citet{2019ApJ...883...99S}.
At  $z\sim$ 9, those data are from \citet{Harikane_2022c}, \citet{2022arXiv220712356D}, \citet{2020MNRAS.493.2059B}, \citet{2019ApJ...883...99S}, 
and \citet{2018ApJ...867..150M}.
At $z\sim$ $10-13$, we take data from \citet{Harikane_2022c}, \citet{2022arXiv220712356D}, \citet{Harikane_2022b}, \citet{2022arXiv220709434N}, \citet{2018ApJ...867..150M}, 
and \citet{2016ApJ...819..129O}.
The fitting forms of the luminosity functions are given by \citet{2022ApJS..259...20H} at $z\simeq 7$, 
\citet{2020MNRAS.493.2059B} and \citet{2013MNRAS.432.2696M} at $z\simeq 8$, and \citet{Harikane_2022c} at $9\lesssim z\lesssim 13$.
}
\label{fig:LFs}
\end{figure*}

@@

Fig.~\ref{fig:LFs} presents our prediction of QLFs at $z>6$ taken from the BH growth model.
We divide the cosmological QLF evolution results into four different epochs; (a) $z\simeq 7$, (b) $z\simeq 8$, (c) $z\simeq 9$, and (d) $z\simeq 10-13$.
The black solid and dashed curves show the QLF for the unobscured population 
and entire population (including unobscured and obscured quasars), respectively.
The obscured fraction based on X-ray samples \citep{2014ApJ...786..104U} is applied to faint objects, 
but the correction factor is highly uncertain at the faint end of $\Muv\gtrsim -20$ mag.
Similar to the BHMF, for all the redshift range, the shape of the QLF is well approximated with a double power-law function,
%
\begin{equation}
\Phi(\Muv) = \frac{\Phi^\ast}
{10^{0.4(\tilde \alpha+1)\Delta \Muv} + 10^{0.4(\tilde \beta+1)\Delta \Muv}},
\end{equation}
%
where $\Phi^\ast$ (in units of Gpc$^{-3}$ mag$^{-1}$) is the overall normalization, $\Muv^\ast$ is the characteristic magnitude,
$\Delta \Muv \equiv \Muv - \Muv^\ast$, and $\tilde \alpha$ is $\tilde \beta$ are the faint and bright-end power-law slopes, respectively.
In Table~\ref{tab:QLF_fit}, we summarize those parameters of the parametric QLFs at $z=6-10$ for the best-fit models.
From higher to lower redshifts, the number density of luminous quasars increases and the bright-end of the distribution rises.

\blue{
We overlay the expected depths and survey volumes of upcoming near-infrared surveys by
Euclid \citep{2011arXiv1110.3193L} and RST \citep{2019arXiv190205569A}.
Here we consider only their wide-area survey layers because other survey programs are not suitable for quasars at $z\gtrsim 7$.
Table~\ref{tab:N_detect} summarizes the coverage and limiting magnitudes in each survey
and lists the expected detection number of quasars.
In priniciple, Euclid will identify quasars up to $z\sim 8$ and discover $O(100-1000)$ quasars at $z\simeq 7-8$.
The RST will potentially be able to explore the very faint ($\Muv>-22$ mag) but abundant quasar population even at $z\gtrsim 10$. 
%in principle, after subtracting galaxy contaminants.
Note that we do not take into account selection criteria to differentiate quasars from other astrophysical sources such
as Galactic brown dwarfs, strong Balmer-break galaxies, and extremely dusty star-forming galaxies at intermediate redshifts.
Therefore, our estimate in Table~\ref{tab:N_detect} gives an upper bound of the number of detectable high-$z$ quasars.
Nevertheless, this simplified assumption yields $\approx 3-7$ times higher values than those predicted by \citet{2019A&A...631A..85E}.
%Studies that take into account contamination from MLT dwarfs and compact early-type galaxies and quasar selction method
%have confirmed the capability of these surveys in searching high-$z$ quasars \citep{2019A&A...631A..85E,2019BAAS...51c.121F}.
%With the simplified assumption of $100\%$ selection function against contaminants,
%our predicted yield of high-$z$ quasars for the Euclid wide survey is a few times higher than that by \citet{2019A&A...631A..85E}.
%Therefore, our estimation should represent an upper bound of the potentially detectable quasar population.
}
% me:On the other hand, we the selection effect to miss some quasars, especially the faint ones. Our prediction is

We also show the galaxy UV luminosity function in each panel
along with the fitting functions in a Schechter (dotted) and double power law (dashed-dotted) shape extended to the brighter end
\citep{2013MNRAS.432.2696M, 2016ApJ...819..129O, 2018ApJ...867..150M, 2019ApJ...883...99S, 2020MNRAS.493.2059B,
2021AJ....162...47B, 2022arXiv220712356D,
2022ApJS..259...20H, Harikane_2022b, Harikane_2022c, 2022arXiv220709434N}.
Galaxies at $\Muv\gtrsim -23$ mag dominate quasars in number at all the redshifts and make it 
difficult to identify rare quasars among those faint objects.
Since the extrapolated galaxy luminosity function with a Schechter shape declines quickly with the luminosity increasing,
the galaxy abundance becomes comparable to that of quasars at $\Muv \sim -23$ mag ($z=7-8$), $-24$ mag ($z=9$), and $-25$ mag ($z=10-13$).
On the other hand, if the bright end of the galaxy luminosity function is extended, the critical UV magnitude shifts to
$\Muv\sim-26$, $-25$, and $\ll -30$ mag for each redshift range.
Future deep and wide surveys will unveil the relative contribution of the two populations and impacts on cosmic reionization
(see also \citetalias{2018ApJ...869..150M}, \citealt{2022NatAs...6..850J}).

@@

The QLF at $z=6$ is shown in the right panel of Fig.~\ref{fig:tau}.
%In the conversion from BHMF to QLF, the ERDF is the same for different $\tlife$ values,
%following the $P(\lambda)$ in Eq.~(\ref{eq:Pl}) with $\lambda_0=0.87$ and $\alpha=0.20$.
Overall, the dependence of the QLF on the choice of $\tau$ shows the same trend as in the BHMF,
%the deviation from the observed QLF of \citetalias{2018ApJ...869..150M} shows similar trends as in the BHMF.
i.e., higher (lower) $\tlife$ values lead to under(over)-production of the luminous quasar population.
The best-fit value of $\tlife (\simeq 18$ Myr) favors a moderate fraction of $t_{\rm S}$.
This result demonstrates the importance of early BHs experiencing multiple accretion episodes to 
reproduce the observed distributions of BH mass and luminosity of quasars at $z\sim$ 6.
% In \S\ref{sec:evol}, we apply this episodic accretion model to individual growth pathways of high-$z$ quasars
% and demonstrate that normal seeding mechanisms can explain the existence of massive BHs with low luminosity
% observed at $z\simeq 6-7$.

>>>>

\shorttitle{BH growth toward $z=$ 6 BHMF \& QLF}
 \shortauthors{Li et al.}
%%
%% You can add a light gray and diagonal water-mark to the first page 
%% with this command:
% \watermark{DRAFT}
%% where "text", e.g. DRAFT, is the text to appear.  If the text is 
%% long you can control the water-mark size with:
%% \setwatermarkfontsize{dimension}
%% where dimension is any recognized LaTeX dimension, e.g. pt, in, etc.
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \graphicspath{{./}{../figs/}}
\graphicspath{{./}{./figs/}}
%% This is the end of the preamble.  Indicate the beginning of the
%% manuscript itself with \begin{document}.

@@

%%%%%%%
%    Fig. 1
%%%%%%%
\begin{figure*}
\centering
\includegraphics[width=160mm]{seedBHMF_z.png}
\caption{
Mass distribution of seed BHs formed in quasar host galaxies at different redshifts ($50\lesssim z \lesssim 10$).
%Each panel includes a redshift span of 10, 
In each panel, the cumulative number distribution of BHs formed prior than the redshift interval is shown with green bars,
and the contribution from newly-born BHs during the interval are highlighted in magenta.
%with the newly formed BHs denoted by the hatched regions, 
%and the cumulative BHs formed previously by the blue region.
The majority of seed BHs form in the epoch of $20\lesssim z \lesssim 40$.
%The most vigorous seed formation epoches are in redshift ranges $z\sim$ $40-30$ and $30-20$.
Overall, the seed BH mass distribution is extended to $M_\bullet \gtrsim 10^5~\Msun$, 
imprinted with various pathways of parent halo assembly and environmental effects as discussed in Section~\ref{sec:seed}.
}
\label{fig:seedmf}
\end{figure*}

@@

We describe how to calculate the time evolution of the BH mass function, 
combining the semi-analytical growth model and ERDF (see Eqs.~\ref{eq:mdot} and \ref{eq:Pl}).
The schematic illustration of our numerical procedure is shown in Fig.~\ref{fig:scheme}. 
We first set a series of checkpoints with an interval of $\tlife$, 
tracing back from $t_{\rm H}(z=6)$ to just prior than the first seed formation time $t_1$, where $t_i$ is the $i$-th seed forming time
as defined in Section~\ref{sec:seed} ($1\leq i\leq N_{\rm tot}$).
Note that the number of the checkpoints is given by $\lceil (t_{\rm H}-t_1)/\tau \rceil+1$, where $\lceil x\rceil \equiv {\rm min}\{n \in \mathbb{Z}~ |~x\leq n\}$.

@@

For the two cases, the best-fit solution consistently shows a small value of $\delta < 0.1$.
This result favors the nearly exponential BH growth model with minor mass dependence (Eqs.~\ref{eq:mdot} and \ref{eq:f_M}).
The typical duration of quasar activity is required to be $\tlife \simeq 20-30$ Myr for both the cases.
This suggests that accreting seed BHs vary their growth speeds and their Eddington ratios 
$\simeq 30-40$ times on average toward the cosmic time at $z\simeq 6$.
The timescale of $\tlife$ is also related to the quasar lifetimes of $t_{\rm Q}\sim 1-10$ Myr, 
which are estimated by the measurement of the physical extents of hydrogen Ly$\alpha$ proximity zones 
observed in the rest-frame UV spectra \citep[e.g.,][]{2018ApJ...867...30E,2019ApJ...884L..19D}.
More detailed comparison is discussed in Section~\ref{sec:evol}.

@@

%%%%%%%
%  Table. 2
%%%%%%%
\begin{deluxetable*}{ccccccccccc}%[t!]
\tablecolumns{1}
\renewcommand\thetable{2} %! fix indexing
\tablewidth{0pt}
\tablecaption{Parametric Quasar Luminosity Function \label{tab:QLF_fit}}
\tablehead{
  & & \multicolumn{4}{c}{Unobscured QLF} && \multicolumn{4}{c}{Intrinsic QLF}  \\
  \cline{3-6} \cline{8-11}
  \colhead{Redshift} && \colhead{$\Phi^\ast$} & \colhead{$\Muv^\ast$} & \colhead{$\tilde \alpha$} & \colhead{$ \tilde \beta$}& &
  \colhead{$\Phi^\ast$} & \colhead{$\Muv^\ast$} & \colhead{$\tilde \alpha$} & \colhead{$\tilde \beta$}\\
  & & \colhead{(Gpc$^{-3}$ mag$^{-1}$)} & \colhead{(mag)} & & & &
  \colhead{(Gpc$^{-3}$ mag$^{-1}$)} & \colhead{(mag)} & &
}
\startdata
$z\sim 6$  && 46.5 & -23.62 & -1.33 & -2.57    && 89.2 & -23.73 & -1.59 & -2.67 \\
$z\sim 7$  && 6.98 & -23.70 & -1.61 & -2.82   && 12.4 & -23.85 & -1.87 & -2.92 \\
$z\sim 8$  && 1.18 & -23.80 & -1.83 & -3.02   && 1.91 & -23.98 & -2.08 & -3.12 \\
$z\sim 9$  && 0.268 & -23.80 & -1.98 & -3.18  && 0.429& -23.96 & -2.23 & -3.28 \\
$z\sim 10$ && 0.0841 & -23.68 & -2.01 & -3.30 && 0.144 & -23.82 & -2.35 & -3.40 \\
\enddata
\tablecomments{Parameters characterizing the double power-law fitting functions
of both the unobscured and intrinsic QLFs.
At $z=6-10$, the QLF evolution is produced by the best-fit BH growth model with $\fseed=0.01$.
}
\end{deluxetable*}

%%%%%%%
%    Table. 3
%%%%%%%
\begin{deluxetable*}{lccccccc}%[t!]
\tablecolumns{1}
\renewcommand\thetable{3} %! fix indexing
\tablewidth{0pt} 
\tablecaption{Detection Number of Quasars Estimated for Future Surveys \label{tab:N_detect}} %! reference indexing
\tablehead{
  \colhead{Telescope} & \colhead{ Survey area} & \colhead{Photo. depth} &  \colhead{$N(z\sim 6)$} & \colhead{$N(z\sim 7)$}& \colhead{$N(z\sim 8)$}&  \colhead{$N(z\sim 9)$}&  \colhead{$N(z\sim 10)$}\\
  \colhead{} & \colhead{(\rm{deg}$^2$)} & \colhead{(AB mag)} & & & & & 
}
\startdata
Wide RST    &  2000  & $YJH$  27   & $6759_{-3185}^{+15} $ &  $1514_{-558}^{+169}$ &  $349_{-131}^{53}$ & $85_{-48}^{+23}$ & $25^{+12}_{-15}$  \\
Wide RST    &  2000  & $\Muv=-26$ & $21_{-5}^{+13} $      &  $2_{-7}^{+8}$        &  --  & -- & --  \\
\hline
Wide Euclid & 15000  & $YJH$  24   & $6654^{+95}_{-2282}$  &  $708^{+213}_{-237}$ &   $93^{+66}_{-56}$ & $14^{+32}_{-27}$ & --  \\
Wide Euclid & 15000  & $YJH$  23   & $2480^{+188}_{-614}$  &  $207^{+144}_{-114}$ &   $23^{+27}_{-19}$ & $3^{+12}_{-10}$ & --  \\
Wide Euclid & 15000  & $\Muv=-26$  & $154^{+100}_{-38}$    &  $12^{+59}_{-50}$    &   $1^{+4}_{-3}$    & -- & --  \\
\enddata
\tablecomments{\blue{
Predicted number of quasars by the BH growth model with $\fseed=0.01$ at each redshift in the two surveys, with the assumption of the survey volume with $\Delta z =1$ and 100\% completeness.
The integration of QLF starts from the 5$\sigma$ point source detection limits of the $YJH$ bands in each survey (27 mag for the RST; 24 mag for Euclid), 
and from the bright absolute magnitude with $\Muv<=-26$ mag.
In addition, with more realistic quasar identification capability down to 23 mag in the $YJH$ bands,
\citet{2019Barnett} predicts the yields of quasar numbers as 117 at $7<z<7.5$, 25 at $7.5<z<8.5$, and 2 at $8.5<z<9$.
We set the same detection limit and find our result in good agreement with theirs.}
}
\end{deluxetable*}

@@

%%%%%%%
%    Fig. 7
%%%%%%%
\begin{figure*}[h!]
\centering
\includegraphics[width=85mm]{LF_spreadz7.pdf}\hspace{2mm}
\includegraphics[width=85mm]{LF_spreadz8.pdf}\\\vspace{5mm}
\includegraphics[width=85mm]{LF_spreadz9.pdf}\hspace{2mm}
\includegraphics[width=85mm]{LF_spreadz10.pdf}
\caption{Predicted quasar luminosity functions \blue{based on the BH growth model with $\fseed=0.01$},
as well as depths and volumes of upcoming surveys.
The black solid and dashed curves present the QLF for the unobscured population (e.g., type-I AGNs) 
and the whole population (i.e., type-I and type-II AGNs), and the shaded region presents the $1\sigma$ spread.
The expected depths and volumes of future surveys with Wide RST and Wide Euclid are overlaid with the blue lines,
indicating the detectability of quasars at these redshifts.
Galaxy UV luminosity functions are plotted with the data points (grey) and fitting functions with a Schechter (dotted curve) and double power-law (dash-dotted curve) shape.
The galaxy luminosity function data at $z\sim$ 7 are taken from \citet{2022ApJS..259...20H} and \citet{2021AJ....162...47B}. 
At $z\sim$ 8, the data points are from \citet{2021AJ....162...47B},
\citet{2022arXiv220712356D}, \citet{2013MNRAS.432.2696M}, \citet{2020MNRAS.493.2059B}, and \citet{2019ApJ...883...99S}.
At  $z\sim$ 9, those data are from \citet{Harikane_2022c}, \citet{2022arXiv220712356D}, \citet{2020MNRAS.493.2059B}, \citet{2019ApJ...883...99S}, 
and \citet{2018ApJ...867..150M}.
At $z\sim$ $10-13$, we take data from \citet{Harikane_2022c}, \citet{2022arXiv220712356D}, \citet{Harikane_2022b}, \citet{2022arXiv220709434N}, \citet{2018ApJ...867..150M}, 
and \citet{2016ApJ...819..129O}.
The fitting forms of the luminosity functions are given by \citet{2022ApJS..259...20H} at $z\simeq 7$, 
\citet{2020MNRAS.493.2059B} and \citet{2013MNRAS.432.2696M} at $z\simeq 8$, and \citet{Harikane_2022c} at $9\lesssim z\lesssim 13$.
}
\label{fig:LFs}
\end{figure*}

@@

Fig.~\ref{fig:LFs} presents our prediction of QLFs at $z>6$ taken from the BH growth model \blue{with $\fseed=0.01$.}
We divide the cosmological QLF evolution results into four different epochs; (a) $z\simeq 7$, (b) $z\simeq 8$, (c) $z\simeq 9$, and (d) $z\simeq 10-13$.
The black solid and dashed curves show the QLF for the unobscured population 
and entire population (including unobscured and obscured quasars), respectively.
The obscured fraction based on X-ray samples \citep{2014ApJ...786..104U} is applied to faint objects, 
but the correction factor is highly uncertain at the faint end of $\Muv\gtrsim -20$ mag.
Similar to the BHMF, for all the redshift range, the shape of the QLF is well approximated with a double power-law function,
%
\begin{equation}
\Phi(\Muv) = \frac{\Phi^\ast}
{10^{0.4(\tilde \alpha+1)\Delta \Muv} + 10^{0.4(\tilde \beta+1)\Delta \Muv}},
\end{equation}
%
where $\Phi^\ast$ (in units of Gpc$^{-3}$ mag$^{-1}$) is the overall normalization, $\Muv^\ast$ is the characteristic magnitude,
$\Delta \Muv \equiv \Muv - \Muv^\ast$, and $\tilde \alpha$ is $\tilde \beta$ are the faint and bright-end power-law slopes, respectively.
In Table~\ref{tab:QLF_fit}, we summarize those parameters of the parametric QLFs at $z=6-10$ for the best-fit models.
From higher to lower redshifts, the number density of luminous quasars increases and the bright-end of the distribution rises.

\blue{
We overlay the expected depths and survey volumes of upcoming near-infrared surveys by
Euclid (\citealt{2011arXiv1110.3193L}; with realistic depth updated in \citealt{2019Barnett}) and RST \citep{2019arXiv190205569A}.
Here we consider only their wide-area survey layers because other survey programs are not suitable for quasars at $z\gtrsim 7$.
Table~\ref{tab:N_detect} summarizes the survey coverage and lists the expected detection number of quasars with varying limiting magnitudes.
We first integrate the QLFs starting from the 5$\sigma$ point source detection limits (27 mag for the RST and 24 mag for Euclid).
In priniciple, Euclid will identify quasars up to $z\sim 8$ and discover $O(100-1000)$ quasars at $z\simeq 7-8$.
The RST will potentially be able to explore the very faint ($\Muv>-22$ mag) but abundant quasar population even at $z\gtrsim 10$. 
Note that here we do not take into account selection criteria to differentiate quasars from other astrophysical sources such
as Galactic brown dwarfs, strong Balmer-break galaxies, and extremely dusty star-forming galaxies at intermediate redshifts.
Therefore, this estimate gives an upper bound of the number of detectable high-$z$ quasars.
Considering 23 mag as the effective depth based on analysis of quasar selection methods for Euclid \citep{2019Barnett},
we find our estimate of the quasars yields in good agreement with their results.
Besides, luminous quasars with $\Muv<=-26$ are predicted to be marginally detected by the RST and Euclid up to $z\sim 7$ and 8, respectively.
However, \citet{2019BAAS...51c.121F} propose that quasar detection can reach $z\sim 9$ by the two surveys,
where they extrapolate the number density of those bright quasars with a milder decay up to higher redshifts. (wli: they used $-0.78$, but we discuss it later with Fig.8.)
The difference of predictions is to be tested by upcoming high-$z$ quasar observations.
%Studies that take into account contamination from MLT dwarfs and compact early-type galaxies and quasar selction method
%have confirmed the capability of these surveys in searching high-$z$ quasars \citep{2019A&A...631A..85E,2019BAAS...51c.121F}.
%With the simplified assumption of $100\%$ selection function against contaminants,
%our predicted yield of high-$z$ quasars for the Euclid wide survey is a few times higher than that by \citet{2019A&A...631A..85E}.
%Therefore, our estimation should represent an upper bound of the potentially detectable quasar population.
}
% me:On the other hand, we the selection effect to miss some quasars, especially the faint ones. Our prediction is

We also show the galaxy UV luminosity function in each panel
along with the fitting functions in a Schechter (dotted) and double powe-law (dashed-dotted) shape extended to the brighter end
\citep{2013MNRAS.432.2696M, 2016ApJ...819..129O, 2018ApJ...867..150M, 2019ApJ...883...99S, 2020MNRAS.493.2059B,
2021AJ....162...47B, 2022arXiv220712356D,
2022ApJS..259...20H, Harikane_2022b, Harikane_2022c, 2022arXiv220709434N}.
Galaxies at $\Muv\gtrsim -23$ mag dominate quasars in number at all the redshifts and make it 
difficult to identify rare quasars among those faint objects.
Since the extrapolated galaxy luminosity function with a Schechter shape declines quickly with the luminosity increasing,
the galaxy abundance becomes comparable to that of quasars at $\Muv \sim -23$ mag ($z=7-8$), $-24$ mag ($z=9$), and $-25$ mag ($z=10-13$).
On the other hand, if the bright end of the galaxy luminosity function is extended, the critical UV magnitude shifts to
$\Muv\sim-26$, $-25$, and $\ll -30$ mag for each redshift range.
Future deep and wide surveys will unveil the relative contribution of the two populations and impacts on cosmic reionization
(see also \citetalias{2018ApJ...869..150M}, \citealt{2022NatAs...6..850J}).

@@

The QLF at $z=6$ is shown in the right panel of Fig.~\ref{fig:tau}.
%In the conversion from BHMF to QLF, the ERDF is the same for different $\tlife$ values,
%following the $P(\lambda)$ in Eq.~(\ref{eq:Pl}) with $\lambda_0=0.87$ and $\alpha=0.20$.
Overall, the dependence of the QLF on the choice of $\tau$ shows the same trend as in the BHMF,
%the deviation from the observed QLF of \citetalias{2018ApJ...869..150M} shows similar trends as in the BHMF.
i.e., higher (lower) $\tlife$ values lead to under(over)-production of the luminous quasar population.
The best-fit value of $\tlife (\simeq 18$ Myr) favors a moderate fraction of $t_{\rm S}$.
This result demonstrates the importance of early BHs experiencing multiple accretion episodes to 
reproduce the observed distributions of BH mass and luminosity of quasars at $z\sim$ 6.
% In Section~\ref{sec:evol}, we apply this episodic accretion model to individual growth pathways of high-$z$ quasars
% and demonstrate that normal seeding mechanisms can explain the existence of massive BHs with low luminosity
% observed at $z\simeq 6-7$.
